---
title: "Ma_ALY6000Project2"
author: "Zihan Ma"
date: "01/22/2023"
output: html_document
---

<!--
Title
-->

<P><BR>
<CENTER>
<FONT SIZE = 4, COLOR="#002FA7">
<B>ALY6000 Introduction to Analytics</B>
<BR><B>Northeastern University</B>
<BR>Zihan Ma
<BR>Date: `r format(Sys.time(), "%m - %d - %Y")` <!--or the time will show in chinese-->
<P>Project Report
</FONT>
</CENTER>

<!--
Library Data
-->
<P><BR>

```{r libraries data, message=FALSE, warning=FALSE}
# Libraries used in this markdown
library(rmarkdown)
library(readxl)
library(readr)
library(tidyverse)
library(dplyr)
library(knitr)
library(kableExtra)
library(magrittr)
library(RColorBrewer)

# Data imported from M2Project_Data.xlsx
library(readxl)
M2Data <- read_excel("DataSets/M2Project_Data-1.xlsx", 
    sheet = "M2Project_V2")
```

<!--
Introduction
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Introduction
</B></FONT>

<P>
<ol>
  <li>The importance of descriptive and inferential statistics</li>
  <P>
  On Bradley University's web page, descriptive statistics is
  included all the population, which means it can minimum the error compare to the 
  reality.
  <P>
  The inferential statistics is doing much better in uncountable, large number population.
  Which we hare or cannot to use descriptive statistics on it.
  <li>Importance of proper data presentation</li>
  <P>
  In "Statistical relevance—relevant statistics, part II: presenting experimental data" 
  by Bernd Klaus, he explained that the proper data presentation will strength the 
  evidence of the analysis and publications. They also helps people find out the patterns 
  and difference in your data set.
  <li>Mention some practical applications of R in data analysis</li>
  <P>
  It easy to find the article that used R to make the chart on Google scholar. For example
, the authors use R language to analyse their data in "An Empirical Data Analytics and Visualization for UBER Services: A Data Analysis Based Web Search Engine". And they used
half of their introduction space to explain the advantage to use R for their analyse tool.
</ol>

<!--
Task 1
Gather total 10 record from data set.
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 1
</B></FONT>
<B>   Present the first 5 and the last 5 records from the dataset</B>

```{r}

first5m2 = head(M2Data, n = 5)
last5m2 = tail(M2Data, n = 5)

firstandlast5m2=rbind(first5m2, last5m2)

kbl(firstandlast5m2) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

<P>
In the ten observations here, we can see there are huge difference in profits, 
and there isn't any obvious rule for these differences, I'm guessing a profit
/sales chart could helpful for find out the parents, or we have to need to increase the sample number.


<!--
Task 2
Give out Market's categories' frequencies
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 2
</B></FONT>
<B>   Present a table with all categories of Market and their frequencies</B>

```{r}

AllMarket = M2Data[, c('Market')]

AllMarketFrequency = table(AllMarket)

kbl(AllMarketFrequency) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```

<P>
The chart here shows the Africa's frequency is clearly lower than other area's frequencies.
It could be affect by Africa's economics conduction deeply, which data that need is not include in this data set.


<!--
Task 3
Give out a graph that shows Market's categories' frequencies
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 3
</B></FONT>
<B>   Present a bar graph with all categories of Market and their frequencies</B>


```{r}
plotAllMarketFrequency = barplot(AllMarketFrequency[order(AllMarketFrequency,decreasing = TRUE)], 
                 main="Plot of All Market's Frequency",
                 xlab="Market Frequency",
                 ylab = "Market Name",
                 border = "red",
                 las = 1,
                 cex.axis = 0.6,
                 cex.names = 0.6,
                 horiz = T,
                 legend.text = rownames(AllMarketFrequency),
                 space = 0.5,
                 xlim = c(0,400),
                 col = brewer.pal(8,"Set1")
                 )
text(AllMarketFrequency[order(AllMarketFrequency,decreasing = TRUE)], 
     plotAllMarketFrequency, 
     AllMarketFrequency[order(AllMarketFrequency,decreasing = TRUE)], 
     cex=0.8, 
     pos=4)

```
<P>
There are almost equally difference between different area frequencies, that tell
us we need to keep in mind that we have to treats all group of data differently because
they have big number off compare to each other.

<!--
Task 4
Use pie chart Market's categories' frequencies for Africa
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 4
</B></FONT>
<B>   Pie chart of product category frequencies in Africa</B>

```{r}

t4Africa = dplyr::filter(M2Data, Market == "Africa")

Product_Categoryt4Africa = t4Africa[, c('Product_Category')]

Product_Categoryt4AfricaFrequency = table(Product_Categoryt4Africa)

pieLabels = paste(unique(sort(t4Africa$Product_Category)),
                  "\n",
                  round(Product_Categoryt4AfricaFrequency,1),
                  "%")

pie(Product_Categoryt4AfricaFrequency,
    labels = pieLabels,
    radius = 0.6,
    col = terrain.colors(3),
    border = "white",
    lty = 1,
    cex=0.8,
    font = 3
    )


legend("bottomleft",
       legend = paste(unique(sort(t4Africa$Product_Category))),
       fill = terrain.colors(3),
       border = "black")

box(col="red")

```

<P>
The plot shows that the major number of companies in Africa is focus on technology area.


<!--
Task 5
Instructor's sample code
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 5
</B></FONT>
<B>   Reproduce codes given by instructor</B>

```{r}

task5_table = table(t4Africa$Product_SubCategory)

t5bar = barplot(task5_table)

text(y=table(t4Africa$Product_SubCategory),
  t5bar,
  table(t4Africa$Product_SubCategory),
  cex=0.8,
  pos=3)


```

<P>
The default plot didn't show all labels properly, it's hard to provide any correct
information from the plot.



<!--
Task 6
Better version of task 5
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 6
</B></FONT>
<B>   Improve bar plot from task 5</B>

```{r}

task6_table = table(t4Africa$Product_SubCategory)

par(mar = c(10,4,0.5,2))


#t6bar = barplot(task6_table)


barplot(task6_table,
main = "Africa Product's Sub Category",
las=2,
# xlab = "Category",
ylab = "Frequency",
col = brewer.pal(9, "Set3"),
ylim=c(0,18),
horiz = FALSE)


text(y=table(t4Africa$Product_SubCategory),
  t5bar,
  table(t4Africa$Product_SubCategory),
  cex=0.8,
  pos=3)


```


<P>
There are many companies are tiring to sell phones in Africa. Which is technology
relative product.



<!--
Task 7
African Market's mean sales for each subcategory.
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 7
</B></FONT>
<B>   Mean sales per subcategory in the African Market</B>

```{r}

meanSales = tapply(t4Africa$Price, t4Africa$Product_SubCategory, mean)

dotchart(meanSales)

```

<P>
For the data set “iris”, the means are setosa: 1.462, versicolor: 4.260, virginica: 5.552.
<P>
We can see the phones have much higher mean number compare to other subcategory in the African Market. It could explain that why so much compines is selling phones in Africa.

<!--
Task 8
The total sales of African Market for each Region 
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 8
</B></FONT>
<B>   Total sales per Region in the African Market</B>

```{r}

sumSales = tapply(t4Africa$Price, t4Africa$Product_SubCategory, sum)

par(mar = c(8,4,0.5,2))

barplot(sumSales,
main = "Africa Product's Total Sales",
las=2,
# xlab = "Sum of Sales",
ylab = "Sum of Sales",
col = brewer.pal(9, "Set3"),
cex.axis = 0.8,
ylim=c(0,45000),
horiz = FALSE)


text(y=table(t4Africa$Product_SubCategory),
  t5bar,
  table(t4Africa$Product_SubCategory),
  cex=1,
  pos=3)

```

<P>
The total sales plot of Africa is very same to Frequency plot we have above in Task 6.
That proved these two group could strongly related to each other.



<!--
Task 9
African Market's average shipping costs for each region
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 9
</B></FONT>
<B>   Mean shipping costs per Region in the African Market</B>

```{r}
meanShipping = tapply(t4Africa$ShippingCost, t4Africa$Product_SubCategory, mean)

par(mar = c(8,4,0.5,2))

barplot(meanShipping,
main = "Africa Product's Average Shipping",
las=2,
# xlab = "Average Shipping",
ylab = "Average Shipping",
col = brewer.pal(9, "Set3"),
cex.axis = 0.8,
ylim=c(0,500),
horiz = FALSE)


text(y=table(t4Africa$Product_SubCategory),
  t5bar,
  table(t4Africa$Product_SubCategory),
  cex=1,
  pos=3)

```

<P>
There isn't great difference between different shipping price in Africa. This data set
 might not very valuable for analyics.



<!--
Task 10
Describe what are integer, factor, double are numeric.
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 10
</B></FONT>
<B>   Descriptions of: integer, factor, double, numeric</B>


<P>
As the website "R for Health Data Science" said,
<P>
"Integers are numbers without decimal places (e.g., 1, 2, 3)"
<P>
"Double stands for “Double-precision floating-point” format (e.g., 1.234, 5.67890)."
<P>
"The values factor levels can take are fixed, and it have an order."
<P>
"Numbers are "R usually refers to numbers as numeric (or num)"


<!--
Task 11
Instructor's example code.
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 11
</B></FONT>
<B>   Duplicate codes presented by instructor</B>

```{r}

par(mfcol = c(2, 1),
    mai = c(1, 1, 0.2, 0.4),
    mar = c(4, 4, 0.5, 2))



boxplot(M2Data$Profits,
        horizontal = T,
        col = rainbow(ncol(trees)),
        xlab = "Profits" ,
        notch = TRUE,
        main = "Profits Boxplot For Whole Dataset")

hist(M2Data$Profits,
     main = "Profits Barplot For Whole Dataset",
     las=1,
     ylim=c(0,350),
     col = brewer.pal(12, "Set3"),
     breaks = 50)






```

<P>
From two plot we can see that there isn't too much off between profits. 
Because more then 50% of data in box plot is concentrate at a short range compare to the distribution. In average most compiles are earning money, but not that much.
<P>
There are many off point on the box plot, but from bar plot we can know these off points are all ignoble.



<!--
Task 12
Latin American market'P's profits
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 12
</B></FONT>
<B>   Profits in the Latin American market</B>



```{r}

t12LATAM = dplyr::filter(M2Data, Market == "LATAM")

par(mfcol = c(2, 1),
    mai = c(1, 1, 0.2, 0.4),
    mar = c(4, 4, 0.5, 2))

boxplot(t12LATAM$Profits,
        horizontal = T,
        col = rainbow(ncol(trees)),
        xlab = "Profits" ,
        notch = TRUE,
        main = "Profits Boxplot For LATAM")

hist(t12LATAM$Profits,
     main = "Profits Barplot For LATAM",
     las=1,
     ylim=c(0,20),
     col = brewer.pal(12, "Set3"),
     breaks = 50)


```
<P>
It's easy to notice from box plot, more than 50% companies are earning money in Latin America.
<P>
And there is only three companies are loosing money greatly(>=500).

<!--
Task 13
Latin America's total sales for each region in a table
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 13
</B></FONT>
<B>   Table total sales per region in Latin America</B>

```{r}
t13LATAM =dplyr::filter(M2Data, Market == "LATAM")
sumLATAMSales = tapply(t13LATAM$Price, t13LATAM$Product_SubCategory, sum)

kbl(sumLATAMSales) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed", "responsive"))

```


<P>
Same to the Africa, Latin America people is also spend a lot of money on buying phones.



<!--
Task 14
Use box plot this time
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 14
</B></FONT>
<B>   Box plot for the profits per region in the Latin American market</B>

```{r}

boxplot(Profits~Region,
data=t13LATAM,
main="Boxplots for each region",
xlab="Region",
ylab="Profits",
col="orange",
border="brown"
)


```


<P>
From the 3 regions in the Latin American, we can see that their distribution are very same to each other. That means Latin American market doesn't not have obvious region differences.



<!--
Task 15
Present cumulative frequencies and probabilities using table
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 15
</B></FONT>
<B>   Table with cumulative frequencies and probabilities</B>

```{r}

t15 = M2Data$Product_SubCategory%>%
  table() %>%
  as.data.frame()%>%
  rename(Frequencies = Freq)%>%
  mutate(CumulativeFrequency = cumsum(Frequencies),
         Probability = Frequencies/nrow(M2Data),
         CumulativeProbability = cumsum(Probability))

knitr::kable(t15,
             digits = 2,
             caption = "Table with cumulative frequencies and probabilities for all region's Product subcategory",
             fomat = "html",
             table.attr = "style='width:40%;'",
             align = 'c') %>%
  kable_classic(bootstrap_option = "striped",
                full_width = TRUE,
                position = "center",
                font_size = 12)

```


<P>
Both Cumulative Frequency and Cumulative Probability are adding to the 1000/1.00 at the end.
Which means there isn't missing observations.



<!--
Task 16
Probability's graphs
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 16
</B></FONT>
<B>   Probability graphs</B>

```{r}

AllMarketFrequency = table(AllMarket)

par(mfrow=c(2,2),mar = c(8, 8, 0.5, 2), mai = c(1,1,0.5,0.2))


Product_Category = M2Data[, c('Product_Category')]
Product_CategorFrequency = table(Product_Category)

pieLabels = paste(M2Data$Product_Category,
                  round(100*Product_CategorFrequency/sum(Product_CategorFrequency), 1)
                  , "%")

pie(t15$Frequencies,
    labels = pieLabels,
    main = "Frequency",
    radius = 1.1,
    col = terrain.colors(17),
    border = "black",
    lty = 1,
    cex=0.4,
    font = 3
    )


barplot(t15$CumulativeFrequency,
main = "Cumulative Frequency",
las=2,
names.arg=c(t15$.),
# xlab = "Average Shipping",
ylab = "Frequency",
col = brewer.pal(12, "Set3"),
cex.axis = 0.6,
cex.lab = 0.6,
cex.names = 0.6,
ylim=c(0,1200),
horiz = FALSE)





pie(t15$Probability,
    labels = pieLabels,
    main = "Probability",
    radius = 1.1,
    col = terrain.colors(17),
    border = "black",
    lty = 1,
    cex=0.4,
    font = 3
    )


barplot(t15$CumulativeProbability,
main = "Cumulative Probability",
las=2,
names.arg=c(t15$.),
# xlab = "Average Shipping",
ylab = "Probability",
col = brewer.pal(12, "Set3"),
cex.axis = 0.6,
cex.lab = 0.6,
cex.names = 0.6,
ylim=c(0,1),
horiz = FALSE)




```


<P>
From plots we can see 1/3 of total Product Category have very little effect to the tatal
sample, which have very little area in the pie plot.



<!--
Task 17
Free graph
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Task 17
</B></FONT>
<B>   Free graph</B>

```{r}


Africa = dplyr::filter(M2Data, Market == "Africa")
AsiaPacific = dplyr::filter(M2Data, Market == "Asia Pacific")
Europe = dplyr::filter(M2Data, Market == "Europe")
LATAM = dplyr::filter(M2Data, Market == "LATAM")
USCA = dplyr::filter(M2Data, Market == "USCA")

AfrMarketFrequency = table(Africa$Product_Category)
AsiMarketFrequency = table(AsiaPacific$Product_Category)
EurMarketFrequency = table(Europe$Product_Category)
LATMarketFrequency = table(LATAM$Product_Category)
USCMarketFrequency = table(USCA$Product_Category)




AddedMarketFrequency=rbind(AfrMarketFrequency, AsiMarketFrequency, EurMarketFrequency, LATMarketFrequency, USCMarketFrequency)

barplot(AddedMarketFrequency,
main = "Added Market Frequency",
#names.arg=c(t15$.),
# xlab = "Average Shipping",
ylab = "Frequency",
col = 1:5,
cex.axis = 1,
cex.lab = 1,
cex.names = 1,
ylim=c(0,450),
horiz = FALSE)

legend("topleft", 
       legend = rownames(AddedMarketFrequency), 
       fill = 1:5, ncol = 2,
       cex = 0.75)

```


<P>
Here is the plot that shows the frequency construction for each region for each Product Category. I believe it can shows the data more directlly.



<!--
Conclusions
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Conclusions
</B></FONT>
<B>   Present</B>

<ol>

<li>
<P>
I learned a lot from this project, especially for give out a good looking format for different plots such as box plot and bar plot. Also It's important to learn using dplyr::filter() to extract a part from the source data frame as a table very easily.

<li>
<P>
By observe this project, we can easily knows that what is the most money earning products,
and the product with list risk. From those observations, we can gives out the advice that 
technology or phones is selling well in most region in the worlds wide.
<li>
<P>
If I'm working on write a good report, I will try to get more recent data set, which will
make my report affect less by any timely or random economics events.
</ol>



<!--
Bibliography
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Bibliography
</B></FONT>

<P>
<P>
Descriptive statistics vs. inferential statistics. Bradley University Online. (2022, March 9). Retrieved January 30, 2023, from https://onlinedegrees.bradley.edu/blog/whats-the-difference-between-descriptive-and-inferential-statistics/ 
<P>
<P>
Klaus, B. (2016). Statistical relevance—relevant statistics, part            ii            : Presenting experimental data. The EMBO Journal, 35(16), 1726–1729. https://doi.org/10.15252/embj.201694659 
<P>
<P>
Pius, E. H. and R. (2021, January 15). R for Health Data Science. 2.2 Variable types and why we care. Retrieved January 30, 2023, from https://argoshare.is.ed.ac.uk/healthyr_book/chap02-vartypes.html 
<P>
<P>
Sudar, K.(1970, January 1). An empirical data analytics and visualization for uber services: A data analysis based web search engine: Semantic scholar. 2022 International Conference on Computer Communication and Informatics (ICCCI). Retrieved January 30, 2023, from https://www.semanticscholar.org/paper/An-Empirical-Data-Analytics-and-Visualization-for-A-Sudar-Nagaraj/771b059c9ca9a7799ab2b19cd5ba7a83129aa218 
<!--
Appendix
-->

<P><BR>
<FONT SIZE = 4, COLOR="#A11515"></B>
Appendix
</B></FONT>

<P>
An R Markdown file has been attached to this report. The name of the file is "Ma_ALY6000Project2..Rmd"


